SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations

### 2015-08-21 paper.tex
.

@@

\begin{abstract}
Online algorithms for robot model-free imitation learning (such as DAgger) can learn policies for problems where system dynamics and reward functions are unknown. However, they impose a burden on supervisors to demonstrate labels (control signals) as the robot encounters new states when iteratively executing its new best policy. We introduce the SHIV algorithm (Svm-based reduction in Human InterVention), which reduces this burden by enabling the robot to only request supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator, 2) 6DOF robot arm push-grasping in clutter (simulation with Box2D), and 3) surgical needle insertion.  Results suggest that SHIV can learn policies and reduce operator burden significantly  (by a factor of \adnote{fill in}).%[I hope more! - ken]

@@

\caption{An illustrative example of different active learning techniques. The shaded and empty circles represent different labeled classes $0$ and $1$. The red empty circles represent label $1$, but are encountered on a different distribution then what was trained on. As illustrated, Confidence and Query by Committee methods measure states far away from the decision boundary of the estimator as low risk, however novelty detection methods measure the level of risk as how far a point is from the trained data. Our modified risk method attempts to measure risk both as a function of how far it is from data trained on and the decision boundary of the estimator. }
\label{fig:support_example}
\end{figure*}

\section{Problem Statement}
\subsection{Notation and Background}
We denote by $\mathcal{X}$ the set consisting of \emph{observable states} for a robot task, consisting, for example of
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of \emph{allowable control inputs} for the robot, which can be discrete or continuous in
nature\adnote{avoid filler words like in nature}. The environment is
modeled\adnote{avoid passive voice; sidd always asked me whenever i wrote that "X is done": "done by who? by jesus?"} to have dynamics that are stationary\adnote{why stationary? what does this mean?} and Markovian, such that the the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in \mathcal{U}$,
so that\adnote{this sentence already contained a so that; break long sentences} $p(\bx_{N+1}|\bu_{N},\bx_{N}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{N+1}|\bu_{N}, \bx_N)$.  
When no confusion arises\adnote{what is confusion arising}, we denote the probability density over the initial state also by $p:\mathcal{X}\to
\mathbb{R}$. By a trajectory, we mean \adnote{new par, a trajectory is a finite ...}a finite series of pairs of $T$ states visited and corresponding
control inputs at these states $\tau = (\mathbf{x}_0,\mathbf{u}_1, ...., \mathbf{x}_T,\mathbf{u}_T)$, where $\bx_i\in \mathcal{X}$
and $\bu_i\in \mathcal{U}$ for $i\in \{0, \ldots, T\}$ and some $T\in \mathbb{N}$.  
At times we will also consider trajectories in state space only\adnote{this makes you seem floppy; do we need this? if so maybe denot by something else, like $\bar{tau}$ and call it a full trajectory vs a state trajectory}, in which case a trajectory corresponds
to a sequence of states $\tau = (\bx_0,....,\bx_T)$ with $\bx_i\in\mathcal{X}$ for $i\in \{0, \ldots, T\}$.

By a policy, we mean\adnote{a policy is a function} a function $\pi: \mathcal{X} \to \mathcal{U}$ from states to control inputs. 
We shall\adnote{we search in a space of parametrized policies} in particular consider classes of policies $\pi_{\theta}:\mathcal{X}\to \mathcal{U}$ parameterized by a
parameter $\theta\in \mathbb{R}^d$ corresponding to support vectors of a Support Vector Machine (SVM).\adnote{this is not a requirement that it's the svm necessarily, can be whatever estimator}
Any such policy, $\pi_{\theta}$ \adnote{missing comma here?} in an environment with probabilistic initial state density and Markovian dynamics
induces a density on state-space trajectories of length $T+1$: $p(\tau | \theta)=
p(\bx_0)\prod_{i=0}^{T-1}p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$, for $\tau = (\bx_0, \bx_1, \ldots, \bx_N)\in
\mathcal{X}^{N+1}$.

\adnote{Let} $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ for $t-1$ time steps\adnote{break sentence. This can be computed by marginalizing ... (explain in words first)} which can be calculated by marginalization $p(\bx_t|\theta) =
\int_{\bx_{t-1}}...\int_{\bx_1} p((\bx_t,...,\bx_1)|\theta) d\bx_{t-1}...d\bx_1$. Following \cite{ross2010reduction},
the average density on states is now defined as $p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)$.
While we do not assume analytic knowledge of the distributions corresponding to: $p(\bx_{t+1}|\bx_t,\bu_t)$, $p(\bx_0)$, $p(\bx_{\bx}|
\bx)$ or $p(\bx|\theta)$, but\adnote{while.., but.. doesn't make sense; "We do not assume... Instead, we assume that..."} instead assume that we have a stochastic real robot or a simulator such that for any state
$\bx_t$ and control $\bu_t$, we can sample the $\bx_{t+1}$ such that\adnote{redundant; we can sample b from $p(b......$"} $\bx_{t+1}$ follows the distribution corresponding
to the density $p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$. In particular\adnote{this is not really in particular; I think you should just say therefore, when rolling out..}, when 'rolling out' trajectories under a policy
$\pi_{\theta}$, we utilize the robot or simulator to 'sample' the resulting stochastic trajectories rather than
estimating $p(\bx|\theta)$ itself.

Typically, the objective in policy learning can be framed\adnote{can be framed by great justice; suggest: typically, the objective of policy learning is to find a policy that minimizes some known cost function:} as minimizing some cost function $C(\tau) = \sum^T_{t=1} c(\bx_t,\bu_t)$
of a trajectory $\tau = (\mathbf{x}_0,\mathbf{u}_1, ...., \mathbf{x}_T,\mathbf{u}_T)\in (\mathcal{X}\times
\mathcal{U})^{T+1}$. The cost function $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in task of inserting a peg into a hole, the distance between the peg's current and desired final state can
be considered \cite{levine2015end}.  In our present setting\adnote{in lfd, the robot does not have access to the cost function itself. Instead, it only has access to a ...}, we do not know the cost function itself, but instead have access to a "supervisor", 
an algorithm or human\adnote{don't start the enumeration with algoirhtm and put human second in a paper motivated by reducing supervisory load, which is more important with human supervision} that we assume uses some near-optimal policy $\tilde{\pi}$ to minimize $C(\tau)$, and we are given
an initial set of $N$ stochastic demonstration trajectories $\mathcal{D} = \lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. \adnote{sentences too long, break them up}

Next, we define a ``surrogate'' loss function $l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

@@

 \subsection{DAgger: Dataset Aggregation}
Note that Eq. \ref{eq:LFD_obj} involves both the parameterized policy $\pi_{\theta}$ and the average distribution of
states with respect to the policy, $p(\bx|\theta)$. Let us recall that Dataset Aggregation (DAgger) is an iterative algorithm developed by \cite{ross2010reduction} which iterates two basic steps for $K$ iterations to find an optimal policy under this objective. 
DAgger iteratively minimizes $\pi_{\theta}$ on the collected dataset of examples $\mathcal{D}$ and then adds new states
of high probability under $\pi_\theta$ by sampling via policy `roll outs' from $p(\bx|\theta)$. 
{\color{blue} [-- F: DAgger and \acro~formal algorithms side by side would be better, also the Step 1/2 below can be
made more precise then and refer to the algorithm --]}

\subsubsection{Step 1}
At iteration $k=0$ an initial dataset $\mathcal{D}_0$ of $N_0$ supervisor demonstration trajectories 
with $T+1$ time steps: $\tau^i=(\bx_0^i, \bu_0^i, \ldots, \bx_T^i, \bu_T^i)\in (\mathcal{X}\times\mathcal{U})^{T+1}$ for $i\in \{0, \ldots, N_0\}$
mapping states to control
inputs is provided. DAgger solves the following optimization problem with respect to the surrogate loss function:

@@

This problem can be considered as a supervised learning problem in particular and can be solved using standard machine
learning techniques, for example a support vector machine or a neural network.

To handle the fact that the supervisor's policy can be noisy, a zero-mean noise term $\epsilon$ 
{ \color{blue} [-- F: Make precise --]}
can be considered as present in policy's output. We employ a regularization technique in the optimization which is
used to control the smoothness of the function that is fit to the sampled data. In practice this regularization corresponds to a penalty term on either the L2 norm on the weights for regression based techniques or the slack coefficient for support vector machines \cite{scholkopf2002learning}.

 \subsubsection{Step 2}
DAgger rolls out the policy $\pi_{\theta_{k=1}}$ to sample states that are likely to occur under $\theta_{k=1}$. For every state visited, DAgger requests the supervisor to provide the appropriate control/label. Formally, for a given sampled trajectory  $\tau = (\bx_0,\bu_0,...,\bx_T,\bu_T )$. The supervisor provides labels $\tilde{\bu}_t$, where $\bu_t \sim \pi(\bx_t) + \epsilon$ for $t\in \{0, \ldots, T\}$.
The states and labeled controls are then aggregated into the next data set of demonstrations $\mathcal{D}_1$.

Steps 1 and 2 are repeated for $K$ iterations or until the cumulative surrogate loss approaches zero. Since DAgger
requires rolling out the policy at each iteration, one way to terminate the algorithm is when the performance of the
rolled out policy matches the intended behavior, or more formally when the surrogate loss is below some predefined
threshold $\gamma$ \footnote{In the orginal DAgger the policy rolled out was stochastically mixed with the supervisor, thus with probability $\beta$ it would either take the supervisor's action or the robots. The use of this stochasitically mix policy was for theoretical analysis. In practice, it is recommended to set $\beta = 0$ to avoid biasing the sampling~\cite{NIPS2014_5421,ross2010reduction}}

Dagger has been experimentally shown to perform  well in both simulation and real world experiments and furthermore under the assumption of the policy being no regret has theoretical properties of being linearly in time horizon T to the supervisor's policy. 
However, policy rollouts can result in the robot entering dangerous states. For example, when DAgger was used on a
quad-copter the robot repeatedly crashed into trees before successfully learning to avoid them \cite{ross2013learning}.
We work under the assumption that a state can be risky for 2 reasons: 1) it lies in an area with a low density of
previously trained states, which potentially means our policy will mis-predict the supervisor and incur high surrogate
loss \cite{tokdar2010importance} 2) the surrogate loss of the current policy at the current state is high, so the state, although visited, does
not model the supervisor's control inputs correctly. Furthermore providing correct control inputs or ``labeling'' at each
iteration and each state in DAgger can be tedious or costly.

To address these limitations, we propose to 1) avoid highly risky states entirely by early termination the policy
roll-out that enter these regions and 2) to recognize non-risky states and to not request human input for these,
thereby reducing the burden on the human trainer.

One way to do this is to define risk in terms of distance to a nearest neighbor, but this approach is vulnerable to
outliers in the known set of states. Another option is to fit a density to the known set of states, which requires an
exponential number of samples in the dimension of the state space \cite{nadaraya1964estimating}. We propose using a
modified version of the technique known as the One Class SVM that implicitly approximately estimates a boundary of a user defined quantile of
a density representing the training data in $\mathcal{X}$ \cite{scholkopf2001estimating}.

\subsection{Estimation of Quantile Level Sets}\label{sec:level}
Support estimation can be defined as minimizing a function that contains the volume of a given probability distribution. Formally we define it as follows let $\mathbf{x}_1,...,\mathbf{x}_n $ be i.i.d. random variables in a set $\mathcal{X}$ with distribution $P$. Let $\mathcal{G}$ be a class of measurable subsets of $\mathcal{X}$ and let $\lambda$ be a real function defined on $\mathcal{G}$. The quantile function with respect to $(P,\lambda,\mathcal{G})$ is

\vspace{-2ex}
\begin{align}\label{eq:quantile}
U(\gamma) = \mbox{inf} \lbrace \lambda(G):P(G) \geq \gamma, G \in \mathcal{G} \rbrace \: 0<\gamma \leq 1
\end{align}

We denote by $G(\gamma)$ the $G \in \mathcal{G}$ that attains the infinimum. The most common choice of $\lambda$ is Lebesgue measure, in which case $G(\gamma)$ is the minimum volume $G \in \mathcal{G}$ that contains at least a fraction $\gamma$ of the the probability mass. Thus, $G(1)$ is the support of the density $p$ corresponding to $P$, assuming it exists.

To handle high dimensional densities $P$, work by Scholk{\"o}pf et al.  looked at representing the class $\mathcal{G}$ via a kernel $k$ as the set of half-spaces in the support vector (SV) feature space \cite{scholkopf2001estimating}. They then minimize a SV  regularizer which, using a kernel, controls the smoothness of the estimated function describing $G$. In terms of the quantile function described in Eq. \ref{eq:quantile}, the approach can be thought of as employing $\lambda(G) = ||w||^2$, where $G_w = \lbrace x: f_w(x) \geq \rho \rbrace$. Here $(w,\rho)$ are a weight vector and offset parameterizing a hyperplane in the feature space associated with a kernel.

Let $\Phi$ be a feature map $\mathcal{X} \rightarrow \mathcal{F}$, i.e. a map into an inner product space $\mathcal{F}$ such that the inner product in the image of $\Phi$ can be computed by evaluating some  kernel

\vspace{-2ex}
\begin{align}
k(\bx_0,\bx_1) = \langle\Phi(\bx_0),\Phi(\bx_1)\rangle
\end{align}

such as the Gaussian kernel

\vspace{-2ex}
\begin{align}
k(\bx_0,\bx_1) = e^{-||\bx_0 - \bx_1||^2/c}
\end{align}

The main idea is to solve an optimization problem that separates the origin, we solve the following quadratic problem:

\vspace{-2ex}
\begin{align}\label{eq:primal_sup}
\underset{w\in F, \mathbf{\xi} \in \mR, \rho \in \mR}{\mbox{min}}\: \frac{1}{2}||w||^2+\frac{1}{vn} \sum^n_i \xi_i - \rho\\
\mbox{s.t} \: \langle w,\Phi(x_i) \rangle \geq \rho - \xi_i, \: \xi_i \geq 0 \notag
\end{align}

The parameter $\nu$ controls the penalty on slack terms and asymptotically approaches is equivalent to $\gamma$ \cite{vert2006consistency}.  The decision function then becomes

\vspace{-2ex}
\begin{align}\label{eq:decision_func}
g(x) = \mbox{sgn}(\langle w,\Phi(x) \rangle-\rho)
\end{align}

Where $g(x) = 1$ indicates in the support region and $(g) = -1$ denotes being outside the support region. The dual form of the optimization takes the form of a Quadratic Program with respect to the Gram Matrix \cite{scholkopf2001estimating}.

Estimating the level set of a distribution can be a measure of the generalization ability of $\pi_{\theta}$ to points not within $\mathcal{D}$. However, it does not account for the fact that for a given $\theta$ or  solution to Eq. \ref{eq:super_objj}. $\pi_\theta$ may not correctly match the supervisor's controls for all points train on. To account for this, we propose a modification to the One Class SVM denote:

\begin{align}
y_i = \left\{
     \begin{array}{lr}
       1 & : \pi_{\theta}(\bx_i) =\bu_i\\
       -1 & : \pi_{\theta}(\bx_i) \neq \bu_i
     \end{array}
   \right.
\end{align}

or $y_i$ is $1$ if the control applied matches the supervisor and $-1$ if it differs. In the regression case the decision between $1$ and $-1$ can be set by a some threshold on a L2-norm or $||\pi_{\theta}(\bx_i)-\bu_i||$. We use $y_i$ to modify the optimization as follows:

\vspace{-2ex}
\begin{align}\label{eq:primal_sup}
\underset{w\in F, \mathbf{\xi} \in \mR^l, \rho \in \mR}{\mbox{min}}\: \frac{1}{2}||w||^2+\frac{1}{vn} \sum^n_i \xi_i - \rho\\
\mbox{s.t} \: y_i \langle w,\Phi(x_i)\rangle \geq \rho - \xi_i, \: \xi_i \geq 0 \label{eq:ineq}
\end{align}

Points that are mis-predicted are then enforced to lie on the opposite of the hyper-plane in the feature space $\phi(\bx)$, which carves out holes in the support region around them. The dual formulation, which is what is solved in practice, can be derived as

\vspace{-2ex}
\begin{align}\label{eq:dual_sup}
\underset{\alpha\in \mathcal{R}}{\mbox{min}} \sum_i^n \sum_j^n \alpha_i\alpha_j y_i y_jk(\bx_i,\bx_j)\\
\mbox{s.t} \: 0 \leq \alpha \leq \frac{1}{\nu n} \:, \sum_i^n \alpha_i = 1 
\end{align}

Our decision boundary function $g(\bx)$ is then derived in terms of the support vectors as $g(\bx) = \sum_i^n \alpha_i y_i k(\bx_i,\bx) - \rho$. To compute the $\rho$ term,we leverage the fact that for an optimal solution the inequalities in Eq. \ref{eq:ineq} are equal for any $\alpha_j$ such that $0 < \alpha_j < \frac{1}{\nu n}$. Therefore we can compute $\rho$ as  $\rho = \sum_i^n y_i \alpha_i k(\bx_i,\bx_j)$ \cite{scholkopf2001estimating}.

\subsection{SHIV: SVM-Based Reduction in Human Intervention}

>>>>

\newcommand{\fpnote}[1]{\ifthenelse{\boolean{include-notes}}%
 {\textcolor{blue}{\textbf{FP: #1}}}{}}

@@

\begin{abstract}
Online algorithms for robot model-free imitation learning (such as DAgger) can learn policies for problems where system
dynamics and reward functions are unknown. However, they impose a burden on supervisors to demonstrate labels (control
signals) as the robot encounters new states when iteratively executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by enabling the robot to only request supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator, 2) 6DOF robot arm push-grasping in clutter (simulation with Box2D), and 3) surgical needle insertion.  Results suggest that SHIV can learn policies and reduce operator burden significantly  (by a factor of \adnote{fill in}).%[I hope more! - ken]

@@

\caption{An illustrative example of different active learning techniques. The shaded and empty circles represent
different labeled classes $0$ and $1$. The red empty circles represent label $1$, but are encountered on a different
distribution then what was trained on
{\color{blue} FP:can we make this more formal sounding? Represent samples from a different distribution?} 
As illustrated, Confidence and Query by Committee methods
measure
{\color{blue} FP:estimate} states far away from the decision boundary of the estimator as having low risk,
however novelty detection methods measure {\color{blue} FP:estimate} the level of risk as
{\color{blue}FP:by} how far a point is from the trained data. Our
modified risk method attempts to measure risk both as a function of how far it
{\color{blue}it gramatically refers to risk,
but you mean sample "it$\mapsto$a sample"?} is from data trained on
{\color{blue}grammatically strange$\mapsto$from the training data} and the decision boundary of the estimator. 
}
\label{fig:support_example}
\end{figure*}

\section{Problem Statement}
\subsection{Notation and Background}
We denote by $\mathcal{X}$ the set consisting of \emph{observable states} for a robot task, consisting, for example of
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of \emph{allowable control inputs} for the robot, which can be discrete or
continuous. We assume that Markovian dynamics, such that the the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$. Under this assumption $p(\bx_{N+1}|\bu_{N},\bx_{N}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{N+1}|\bu_{N}, \bx_N)$.  
We denote the probability density over the initial state also by $p:\mathcal{X}\to
\mathbb{R}$. By a trajectory, we mean a finite series of pairs of $T$ states visited and corresponding
control inputs at these states $\tau = (\mathbf{x}_0,\mathbf{u}_0, ...., \mathbf{x}_T,\mathbf{u}_T)$, where $\bx_i\in \mathcal{X}$
and $\bu_i\in \mathcal{U}$ for $i\in \{0, \ldots, T\}$ and some $T\in \mathbb{N}$.  
For a given trajectory $\tau$ as above, we denote by $\hat{\tau}$ the corresponding trajectory in state space,
$\hat{\tau} = (\bx_0,....,\bx_T)$ with $\bx_i\in\mathcal{X}$ for $i\in \{0, \ldots, T\}$.

A policy is a function $\pi: \mathcal{X} \to \mathcal{U}$ from states to control inputs. 
We consider a space of policies $\pi_{\theta}:\mathcal{X}\to \mathcal{U}$ parameterized by a
parameter $\theta\in \mathbb{R}^d$. Any such policy $\pi_{\theta}$ in an environment with probabilistic initial state density and Markovian dynamics
induces a density on state-space trajectories of length $T+1$: $p(\tau | \theta)=
p(\bx_0)\prod_{i=0}^{T-1}p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$, for $\tau = (\bx_0, \bx_1, \ldots, \bx_N)\in
\mathcal{X}^{N+1}$.

Let $p(\bx_t|\theta)$ denote the value of the density of states visited at time $t$ if the robot follows the policy
$\pi_{\theta}$ for $t-1$ time steps. This density can be computed by marginalization: $p(\bx_t|\theta) =
\int_{\bx_{t-1}}...\int_{\bx_1} p((\bx_t,...,\bx_1)|\theta) d\bx_{t-1}...d\bx_1$. Following \cite{ross2010reduction},
the average density on states is now defined as $p(\bx|\theta) = \frac{1}{T} \sum^T_{t=1} p(\bx_t|\theta)$.
While we do not assume analytic knowledge of the distributions corresponding to: $p(\bx_{t+1}|\bx_t,\bu_t)$, $p(\bx_0)$, $p(\bx_{\bx}|
\bx)$ or $p(\bx|\theta)$, we assume that we have a stochastic real robot or a simulator such that for any state
$\bx_t$ and control $\bu_t$, we can sample the $\bx_{t+1}$ from the density $p(\bx_{t+1}|\pi_{\theta}(\bx_t),\bx_t)$. 
Therefore, when 'rolling out' trajectories under a policy
$\pi_{\theta}$, we utilize the robot or a simulator to sample the resulting stochastic trajectories rather than
estimating $p(\bx|\theta)$ itself.

Typically, the objective of policy learning is to find a policy that minimizes some known cost function
$C(\tau) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\tau = (\mathbf{x}_0,\mathbf{u}_1, ...., \mathbf{x}_T,\mathbf{u}_T)\in (\mathcal{X}\times
\mathcal{U})^{T+1}$. The cost function $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in task of inserting a peg into a hole, the distance between the peg's current and desired final state can
be considered \cite{levine2015end}.  
In our learning from demonstration setup, the robot does not have access to the cost function itself. Instead, it only has access to a
a `supervisor' that we assume uses some near-optimal policy $\tilde{\pi}$ to minimize $C(\tau)$. We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\mathcal{D} = \lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy.

Next, we define a `surrogate' loss function $l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

@@

 \subsection{DAgger: Dataset Aggregation}
Note that Eq. \ref{eq:LFD_obj} involves both the parameterized policy $\pi_{\theta}$ and the average distribution of
states with respect to the policy, $p(\bx|\theta)$. Let us recall that Dataset Aggregation (DAgger) is an iterative algorithm developed by \cite{ross2010reduction} which iterates two basic steps for $K$ iterations to find an optimal policy under this objective. 
DAgger iteratively minimizes $\pi_{\theta}$ on the collected dataset of examples $\mathcal{D}$ and then adds new states
of high probability under $\pi_\theta$ by sampling via policy `roll outs' from $p(\bx|\theta)$. 
\fpnote{DAgger and \acro~formal algorithms side by side would be better, also the Step 1/2 below can be
made more precise then and refer to the algorithm}

\subsubsection{Step 1}
\fpnote{this is really start of iteration `i', as we circle back}
At iteration $k=0$ an initial dataset $\mathcal{D}_0$ of $N_0$ supervisor demonstration trajectories 
with $T+1$ time steps: $\tau^i=(\bx_0^i, \bu_0^i, \ldots, \bx_T^i, \bu_T^i)\in (\mathcal{X}\times\mathcal{U})^{T+1}$ for $i\in \{0, \ldots, N_0\}$
mapping states to control
inputs is provided. DAgger solves the following optimization problem with respect to the surrogate loss function:

@@

This problem can be considered as a supervised learning problem which can be solved using standard machine
learning techniques, for example a support vector machine or a neural network.

To handle the fact that the supervisor's policy can be noisy, a zero-mean noise term $\epsilon$ 
can be considered as present in policy's output. We employ a regularization technique in the optimization which is
used to control the smoothness of the function that is fit to the sampled data. In practice this regularization corresponds to a penalty term on either the L2 norm on the weights for regression based techniques or the slack coefficient for support vector machines \cite{scholkopf2002learning}.

 \subsubsection{Step 2}
DAgger rolls out the policy $\pi_{\theta_{k=1}}$ to sample states that are likely to occur under $\theta_{k=1}$. For every state visited, DAgger requests the supervisor to provide the appropriate control/label. Formally, for a given sampled trajectory  $\tau = (\bx_0,\bu_0,...,\bx_T,\bu_T )$. The supervisor provides labels $\tilde{\bu}_t$, where $\bu_t \sim \pi(\bx_t) + \epsilon$ for $t\in \{0, \ldots, T\}$.
The states and labeled controls are then aggregated into the next data set of demonstrations $\mathcal{D}_1$. 
\fpnote{terms such as aggregated etc would be much clearer if we just refer to a particular line in the algorithm}

Steps 1 and 2 are repeated for $K$ iterations or until 
the cumulative surrogate is smaller than some predefined threshold\footnote{In the orginal DAgger the policy rolled out
was stochastically mixed with the supervisor, thus with probability $\beta$ it would either take the supervisor's action
or the robots. The use of this stochasitically mix policy was for theoretical analysis. In practice, it is recommended
to set $\beta = 0$ to avoid biasing the sampling~\cite{NIPS2014_5421,ross2010reduction}}.

Dagger has been experimentally shown to perform  well in both simulation and real world experiments and furthermore under the assumption of the policy being no regret has theoretical properties of being linearly in time horizon T to the supervisor's policy. 
\fpnote{How about: Dagger has been experimentally shown to perform well in both simulation and real world
    experiments. Under a technical `no regret' assumption, Ross et al.\cite{ross2013learning} furthermore
deduced error bounds on the expected loss which scale linearly in the time horizon $T$.}

However, policy rollouts can in practice result in the robot entering dangerous states. For example, when DAgger was used on a
quad-copter the robot repeatedly crashed into trees before successfully learning to avoid them \cite{ross2013learning}.
We work under the assumption that a state can be risky for 2 reasons: 1) it lies in an area with a low density of
previously trained states, which can cause our policy to mis-predict the supervisor and incur high surrogate
loss \cite{tokdar2010importance} 2) the surrogate loss of the current policy at the current state is high, so that the state, although visited, does
not model the supervisor's control inputs correctly. Furthermore providing correct control inputs or ``labeling'' at each
iteration and each state in DAgger can be tedious or costly.

To address these limitations, we propose to 1) avoid highly risky states entirely by early termination of the policy
roll-outs that enter these regions and 2) to recognize non-risky states and to not request supervisor input for these,
thereby reducing the burden on the human trainer.

One way to do this is to define risk in terms of distance to a nearest neighbor, but this approach is vulnerable to
outliers in the known set of states. Another option is to fit a density to the known set of states, which requires an
exponential number of samples in the dimension of the state space \cite{nadaraya1964estimating}. We propose using a
modified version of the technique known as the One Class SVM that implicitly approximately estimates a boundary of a user defined quantile of
a density representing the training data in $\mathcal{X}$ \cite{scholkopf2001estimating}.
\fpnote{Refer to figure, replace "one way to do"  by "an approach to this problem is to" or similar - less informal}

\subsection{Estimation of Quantile Level Sets}\label{sec:level}
\fpnote{my git commit on this section was overwritten by an old commit at version 10 - reintroduced modifications here,
Michael, please check this is still accurate with recent modifications}
We consider the problem of estimating the quantile level-sets of a distribution $P$ on a set $\mathcal{X}$ by means of a finite set of
independent and identically distributed samples $\mathbf{x}_1,...,\mathbf{x}_n\in \mathcal{X}$.
In most general terms, the quantile function for $P$ and subject to a class of measurable subsets $\mathcal{G}$ of $\mathcal{X}$ is
defined by
\begin{align}\label{eq:quantile}
U(\gamma) = \mbox{inf} \lbrace \lambda(G):P(G) \geq \gamma, G \in \mathcal{G} \rbrace \: 0<\gamma \leq 1
\end{align} 
$\lambda:\mathcal{G}\to \mathbb{R}$ above denotes a volume measure which most commonly is given by the Lebesgue measure.
Suppose furthermore that $G:[0,1]\to \mathcal{G}$ assigns a set $G(\gamma) \in \mathcal{G}$ that attains the infinum
measure (i.e. volume) for each $\gamma\in [0,1]$ (this set is in general not necessarily unique). 
$G(\gamma)$ denotes a set of minimum measure $G \in \mathcal{G}$ with $P(G(\gamma))\ge \gamma$. Note in particular that $G(1)$ is the support of the density $p$ corresponding to $P$, if $p$ exists.

To handle distributions defined on high-dimensional spaces $\mathcal{X}$, work by Scholk{\"o}pf et al. looked at representing the class $\mathcal{G}$ via a kernel $k$ as the set of half-spaces in the support vector (SV) feature space \cite{scholkopf2001estimating}. 
By minimizing a support vector regularizer controlling the smoothness of the estimated level set function this work
derives an approximation of the quantile function described in Eq. \ref{eq:quantile}, which in particular has rigorous
convergence guarantees asymptotically when a normalized Gaussian kernel $k$ is chosen and bandwidths decay at an
appropriately chosen rate as the number of samples tends to infinity \cite{vert2006consistency}.
This approach can be thought of as employing $\lambda(G) = ||w||^2$, where $G_w = \lbrace x: f_w(x) \geq \rho \rbrace$,
$f_w(\mathbf{x}) = \sum_i w_i k(\mathbf{x}_i, \mathbf{x})$
and $(w,\rho)$ denote the weight vector and offset parameterizing a hyperplane in the feature space associated with a
Gaussian kernel $k(\bx_0,\bx_1) = e^{-||\bx_0 - \bx_1||^2/c}$.

Let $\Phi:\mathcal{X}\to \mathcal{F}$ denote the feature map corresponding to our exponential kernel, mapping the
observation space $\mathcal{X}$ into a Hilbert space $(\mathcal{F}, \langle, \rangle)$ such that $k(\bx, \bx') = \langle
\Phi(\bx), \Phi(\bx')\rangle$.

The One Class SVM proposed by \cite{scholkopf2001estimating} determines a hyperplane in feature space $\mathcal{F}$
maximally separating the input data from the origin:
\vspace{-2ex}
\begin{align}\label{eq:primal_sup}
    \underset{w\in \mathcal{F}, \mathbf{\xi} \in \mathbb{R}, \rho \in \mathbb{R}}{\mbox{min}}\: \frac{1}{2}||w||^2+\frac{1}{vn} \sum^n_i \xi_i - \rho\\
\mbox{s.t} \: \langle w,\Phi(x_i) \rangle \geq \rho - \xi_i, \: \xi_i \geq 0 \notag.
\end{align}

Here, the parameter $\nu$ controls the penalty or `slack term' and asymptotically approaches $\gamma$ \cite{vert2006consistency}
in the quantile definition when bandwidths decay at appropriate rates as the number of samples increases. The decision
function, determining point membership in the approximate quantile levelset is given by

\vspace{-2ex}
\begin{align}\label{eq:decision_func}
g(x) = \mbox{sgn}(\langle w,\Phi(x) \rangle-\rho).
\end{align}

Where, for $x\in \mathcal{X}$, $g(x)=0$ if $x$ lies on the quantile levelset,
$g(x) = 1$ if $x$ is strictly in the interior of the quantile super-levelset and $g(x) = -1$ 
if $x$ lies strictly in the quantile sub-levelset. The dual form of the optimization yields a Quadratic Program 
that can be solved efficiently \cite{scholkopf2001estimating}.

For training data in $\mathcal{X}$, we consider the quantile super-levelset for a user-defined threshold $\mu$ as a natural region within which
a trained policy $\pi_{\theta}$ can likely be learned since a sufficient data-density is available.
However, even when sufficient data is available, the associated control inputs may be inconsistent or noisy and a resulting policy
optimizing Eq. \ref{eq:super_objj} may still incur a large surrogate loss. To account for this, we propose a
modification to the One Class SVM:

\begin{align}
y_i = \left\{
     \begin{array}{lr}
         1 & : l(\pi_{\theta}(\bx_i),\bu_i)\le \varepsilon\\
         -1 & : l(\pi_{\theta}(\bx_i),\bu_i)>\varepsilon
     \end{array}
   \right.
\end{align}
Where, in the case when $l$ denotes discrete $0-1$ loss, we set $\varepsilon = 0$, while in the continuous $L_2$ loss
case, $\varepsilon$ is a user defined threshold specifying allowable surrogate loss.
We use $y_i$ to modify the One Class SVM optimization objective as follows:

\vspace{-2ex}
\begin{align}\label{eq:primal_sup}
    \underset{w\in F, \mathbf{\xi} \in \mathbb{R}^l, \rho \in \mathbb{R}}{\mbox{min}}\: \frac{1}{2}||w||^2+\frac{1}{vn} \sum^n_i \xi_i - \rho\\
\mbox{s.t} \: y_i \langle w,\Phi(x_i)\rangle \geq \rho - \xi_i, \: \xi_i \geq 0 \label{eq:ineq}.
\end{align}
